Circular Regression Trees and Forests with an Application to Probabilistic Wind Direction Forecasting

### 2020-07-05 pkg/circtree/inst/wind_paper/draft/draft.Rnw
Include Achim's newly written abstract
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@3042 255f4249-2263-4417-b1ea-8498779e27a7

Former-commit-id: a3ef4b7b7d9b54c87f20174fa2e674c971566acb

\Abstract{
%SUBMITTED VERSION:
%While circular data occur in a wide range of scientific fields, the methodology
%for distributional modeling and probabilistic forecasting of circular response
%variables is rather limited. Most of the existing methods are built on the
%framework of generalized linear and additive models, which are often
%challenging to optimize and interpret. Therefore, building on previous ideas
%for trees modeling circular means, we suggest a distributional approach for
%regression trees and random forests yielding probabilistic forecasts based on
%the von~Mises distribution. The resulting tree-based models simplify the
%estimation process by using the available covariates for partitioning the data
%into sufficiently homogeneous subgroups so that a simple von~Mises distribution
%without further covariates can be fitted to the circular response in each
%subgroup. These circular regression trees are straightforward to interpret, can
%capture nonlinear effects and interactions, and automatically select the
%relevant covariates that are associated with either location and/or scale
%changes in the von~Mises distribution. Combining an ensemble of circular
%regression trees to a circular regression forest can regularize and smooth the
%covariate effects. The new methods are evaluated in a case study on
%probabilistic wind direction forecasting at two Austrian airports, considering
%other common approaches as a benchmark.
%
%NEW VERSION:
Circular data occur in a wide range of scientific fields, however the
methodology for distributional modeling and probabilistic forecasting of
circular response variables is rather limited. While most of the existing
methods are built within the framework of generalized linear and additive
models, which are often challenging to optimize and interpret, we propose a
distributional approach for regression trees yielding probabilistic forecasts
based on the von~Mises distribution. The resulting tree-based models use the
available covariates for partitioning the data into sufficiently homogeneous
subgroups, allowing for a simple circular distributional fit in each subgroup.
These circular regression trees are straightforward to interpret, can capture
nonlinear effects and interactions, and automatically select the relevant
covariates that are associated with changes in any distribution parameter.
Ensembles or forests of such circular trees can further improve the forecasts
by regularizing and stabilizing the model. The newly developed methodology is
illustrated in an application on short-term probabilistic wind direction
forecasting for Innsbruck and Vienna Airport required to direct airplanes to a
safe landing. The predictive skill of the novel approaches is benchmarked with
an additive regression model plus a persistency and a climatological model,
employing the circular continuous ranked probability score.}

>>>>

\Abstract{
%SUBMITTED VERSION:
%While circular data occur in a wide range of scientific fields, the methodology
%for distributional modeling and probabilistic forecasting of circular response
%variables is rather limited. Most of the existing methods are built on the
%framework of generalized linear and additive models, which are often
%challenging to optimize and interpret. Therefore, building on previous ideas
%for trees modeling circular means, we suggest a distributional approach for
%regression trees and random forests yielding probabilistic forecasts based on
%the von~Mises distribution. The resulting tree-based models simplify the
%estimation process by using the available covariates for partitioning the data
%into sufficiently homogeneous subgroups so that a simple von~Mises distribution
%without further covariates can be fitted to the circular response in each
%subgroup. These circular regression trees are straightforward to interpret, can
%capture nonlinear effects and interactions, and automatically select the
%relevant covariates that are associated with either location and/or scale
%changes in the von~Mises distribution. Combining an ensemble of circular
%regression trees to a circular regression forest can regularize and smooth the
%covariate effects. The new methods are evaluated in a case study on
%probabilistic wind direction forecasting at two Austrian airports, considering
%other common approaches as a benchmark.
%
%NEW LISA AND MORITZ VERSION
%Circular data occur in a wide range of scientific fields, however the
%methodology for distributional modeling and probabilistic forecasting of
%circular response variables is rather limited. While most of the existing
%methods are built within the framework of generalized linear and additive
%models, which are often challenging to optimize and interpret, we propose a
%distributional approach for regression trees yielding probabilistic forecasts
%based on the von~Mises distribution. The resulting tree-based models use the
%available covariates for partitioning the data into sufficiently homogeneous
%subgroups, allowing for a simple circular distributional fit in each subgroup.
%These circular regression trees are straightforward to interpret, can capture
%nonlinear effects and interactions, and automatically select the relevant
%covariates that are associated with changes in any distribution parameter.
%Ensembles or forests of such circular trees can further improve the forecasts
%by regularizing and stabilizing the model. The newly developed methodology is
%illustrated in an application on short-term probabilistic wind direction
%forecasting for Innsbruck and Vienna Airport required to direct airplanes to a
%safe landing. The predictive skill of the novel approaches is benchmarked with
%an additive regression model plus a persistency and a climatological model,
%employing the circular continuous ranked probability score.
While circular data occur in a wide range of scientific fields, the methodology 
for distributional modeling and probabilistic forecasting of circular response
variables is rather limited. Most of the existing methods are built on the
framework of generalized linear and additive models, which are often challenging
to optimize and interpret. Specifically, capturing abrupt changes or
interactions is not straightforward but often relevant, e.g., for modeling wind
directions subject to different wind regimes. Additionally, automatic covariate
selection is desirable when many predictor variables are available, as often the
case in weather forecasting.

To address these challenges we suggest a general distributional approach using
regression trees and random forests to obtain probabilistic forecasts for
circular responses. Using trees simplifies model estimation as
covariates are used only for partitioning the data and subsequently just a
simple von~Mises distribution is fitted in the resulting subgroups.
Circular regression trees are straightforward to interpret, can capture
nonlinear effects and interactions, and automatically select covariates
affecting location and/or scale in the von~Mises distribution. Circular random
forests regularize and smooth the effects from an ensemble of trees.

The new methods are applied to probabilistic wind direction forecasting at two
Austrian airports, considering other common approaches as a benchmark.}
